Renal IV

### 2017-06-01 renal4.rmd
added ryan's farewell

>>>>

> 22/12M child, febrile, vomited, diarrhoea

GIT symptoms from systemic things such as: thyroid problems, infn in urinary tract cytokines (IL-1, TNF-$\alpha$, IL-5), sick from URTI can give diarrhoea.

Pathophysiological processes causing fever: malignancy, infn, inflammation (autoimmune), medications/immunisations,

Pathophys involved with diarrhoea and vomiting: fluid shifts associated with fever, inflamm, infn from {protozoa, bacteria, viral}, toxicity, thyrotoxicosis, pain, overeating, motion sickness, allergic response

NSAIDS and paracetamol make fever abate.

Most common cause of fever in child: URTI such as otitis media, viral, skin infections

Paediatrics is like veterinary science => have to ask random q's since child can't explain things themselves. \textbf{Always ask more and examine more in children}.

Conditions not to miss in a febrile child (potentiall fatal or life altering)

-  Ca
-  Septic shock
-  Meningococcal disease (Neiserria meningitis (gram neg coccus); meningitis): neck stiffness/rigidity, fever, koenig's sign, Brudzinski's sign (more for adults), rash (late sign; caused by septic emboli getting into tissue and causing microinfarctions => non-blanching rash which separates it from dermal rashes), decreased level of consciousness
-  Measles: C's = {koplic spots (white spots in mouth), conjuctivitis, coryza}, fever

Lethargy is basically saying the child is unwell and needs some help.

Fever

-  Duration: $\geq5$ days: have to consider Kawasaki disease which is an autoimmune disease
-  Frequency: cyclical (malaria,...)

Asymptomatic UTIs: no dysuria = {young, elderly}

Take samples **before** starting antibiotic treatment.

Asymptomatic bacteriauria (bacteria in urine) which isn't a problem unless symptoms, pregnant, or transplant.

High fever usually indicates pyelonephritis (also common in first year of life) rather than cystitis.

Ureters implant obliquely on the bladder which is a functional valve (not anatomical).
